---
title: "Key Practices for the Language Scientist"
subtitle: "HW for Session 2b: Power Analysis & \n Session 5: Best Practices for Coding in R"
author: "Monica Wagner"
date: "4/15/2021"
output: pdf_document
---

This code was written for the 2021 Key Practices IMPRS course in order to perform a power analysis based on a simulated data set.

```{r setup, include=FALSE}

## Chunk settings
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

## Libraries
library(here)
library(pwr)
library(simr)
library(tidyverse)

## Workgroup folder
knitr::opts_knit$set(root.dir = here::here())
```

## Assignment

Run a power analysis to find out how much participants you need to get a power of 0.80.

## Analysis of simulated dataset

```{r sim_load, include = FALSE}
## Load data
simData = read.table("SimData_SoPVoc.csv", sep = ",", header = TRUE)
```

```{r sim_plot}
## Plot data
ggplot(simData, aes(x = produce, y = meanRT)) +
  geom_point() +
  geom_smooth(method=lm) +
  labs(title = "Scatterplot of infants' expressive vocabulary and \nword recognition reaction time (RT)", x = "Expressive vocabulary (#words produced)", y = "Word recognition RT (ms)") +
  theme_minimal()
```

```{r sim_corr}
simCorrResults = cor.test(simData$meanRT, simData$produce)
print(simCorrResults)
```

A Pearson's correlation was conducted on the data set between number of words produced and mean word recognition RT. The results revealed a correlation of r = r(`r simCorrResults$parameter`) = `r round(simCorrResults$estimate, digits = 2)`, p = `r round(simCorrResults$p.value, digits = 3)`.

Based on this, we'll assume a small effect size of 0.2 for our envisioned study.

## Power calculation

We run a power analysis for power = 0.80, r = 0.2, and p = 0.05.

```{r}
pwrResults = pwr.r.test(n = NULL,
                        r = 0.2,
                        sig.level = 0.05,
                        power = 0.80,
                        alternative = "two.sided")
print(pwrResults)
```

Based on the power analysis, in order to have an 80% chance of detecting an effect, at least `r floor(pwrResults$n)` participants should be tested.

